### 2009-01-21 docs/flowchart/flowchart.tex
fixed some typos, added more detail about mapping and where it happens in the algorithm

\rho^n & & & \longrightarrow & & & & & \rho^{n+1} \\

@@

(\rho h)^n & \longrightarrow & (\rho h)^{(1)} & \longrightarrow & (\rho h)^{(2a)} & \longrightarrow & (\rho h)^{(2)} & \longrightarrow & (\rho h)^{n+1}\\

@@

Note that this chart indicates, for example, that $\rho^n = \rho^{(1)}$, $\rho^{(2a)} = \rho^{(2)} = \rho^{n+1}$, etc.\\ \\

@@

After the new mass fractions have been computed, the reaction rates are defined as:
\begin{equation}
(\rho\omegadot_k)^{\outp} = \frac{\rho^{\outp} ( X_k^{\outp} - X_k^{\inp})}{\Delta t /2} \enskip,
\end{equation}
and the nuclear energy generation rate is defined as
\begin{equation}
(\rho\Hnuc) = -\sum_k(\rho\omegadot_k)^{\outp} \; q_k \enskip.
\end{equation}
for the case where only strong-mediated reactions are involved.  As
$\Hnuc$ is an output of the reaction network, more general networks
(i.e.\ involving weak interactions) can easily be added.  The enthalpy
update incorporates the external heating, $(\rho\Hext)^{\inp}$, and is
updated by
\begin{equation}
(\rho h)^{\outp} = 
(\rho h)^{\inp} + \frac{\dt}{2} (\rho\Hnuc) + \frac{\dt}{2} (\rho\Hext)^{\inp}\enskip.
\end{equation}

@@

  \item Compute the species edge states, $(\rho X_k)^{(1),\nph,\star}$,
  for the conservative update of $(\rho X_k)$. 
  Here we predict $\rho'$ and $X_k$ to the edges to obtain 
  $\rho^{'(1),\nph,\star}$ and $X_k^{(1),\nph,\star}$ using a second-order
  Taylor expansion in space and time, as described in paper II,
  Appendix A, using $\Vb = \uadvone+w_0^{\nph,\star} \er$.  We use the base
  state density, $\rho_0$, to convert this to an edge state for $\rho$.
  We use 4th-order spatial averaging to get $\rho_0$ on radial faces,
\begin{equation}
\rho^{(1),\nph,\star} = 
\rho^{'(1),\nph,\star} + \frac{\rho_0^n + \rho_0^{(2a),\star}}{2} \enskip ,
\label{eq:rhoXedgestate}
\end{equation}

@@

  \item Compute the enthalpy edge state, $(\rho h)^{(1),\nph,\star},$
    for the conservative update of $(\rho h).$  Here we predict $(\rho
    h)'$ to the edges to obtain $(\rho h)^{'(1),\nph,\star}$, 
    using a second-order Taylor expansion in space
    and time, as described in paper II, Appendix A, using $\Vb =
    \uadvone+w_0^{\nph,\star} \er$.  We do not include the reaction
    terms in the enthalpy prediction, since we accounted for them
    already in {\bf React State}.  We use the base state enthalpy,
    $(\rho h)_0$, to convert this to an edge state for $(\rho h)$.
    We use 4th-order spatial averaging to get $(\rho h)_0$ on radial faces,
\begin{equation}
(\rho h)^{(1),\nph,\star} = 
(\rho h)^{'(1),\nph,\star} + \frac{(\rho h)_0^n + (\rho h)_0^{(2),\star}}{2}
\enskip .
\end{equation}

   We note that for spherical geometry, the edge state of $(\rho h)_0$
   is computed by averaging $h_0 \equiv (\rho h)_0/\rho_0$ and $\rho_0$
   to the edge separately, and then computing 
   ${(\rho h)_0}_{r+\myhalf} = {\rho_0}_{r+\myhalf} {h_0}_{r+\myhalf}$.

@@

\begin{equation}
\rho^{(1),\nph} = \rho^{'(1),\nph} + \frac{\rho_0^n + \rho_0^{(2a)}}{2}\enskip ,
\end{equation}

@@

  \item Compute the enthalpy edge state, $(\rho h)^{(1),\nph}$, for
    the conservative update of $(\rho h)$. Here we predict $(\rho h)'$
    to the edges to obtain $(\rho h)^{'(1),\nph}$, 
    using a second-order Taylor expansion in space and
    time, as described in paper II, Appendix A, using $\Vb =
    \uadvtwo+w_0^{\nph} \er$.  Again, we do not include the reaction
    terms in the enthalpy prediction since we accounted for them
    already in {\bf React State}. We use the base state enthalpy,
    $(\rho h)_0$, to convert these to edge states for $(\rho h)$,
\begin{equation}
(\rho h)^{(1),\nph} = 
(\rho h)^{'(1),\nph} + \frac{(\rho h)_0^n + (\rho h)_0^{(2)}}{2} \enskip .
\end{equation}

   We note that for spherical geometry, the edge state of $(\rho h)_0$
   is computed by averaging $h_0 \equiv (\rho h)_0/\rho_0$ and $\rho_0$
   to the edge separately, and then computing 
   ${(\rho h)_0}_{r+\myhalf} = {\rho_0}_{r+\myhalf} {h_0}_{r+\myhalf}$.
   \item Evolve $(\rho h)^{(1)} \rightarrow (\rho h)^{(1a)}$ without
     explicitly including the reaction terms,

>>>>

\rho^n & = & \rho^{(1)} & \longrightarrow & & & \rho^{(2)} & = & \rho^{n+1} \\

@@

(\rho h)^n & \longrightarrow & (\rho h)^{(1)} & \longrightarrow & (\rho h)^{(1a)} & \longrightarrow & (\rho h)^{(2)} & \longrightarrow & (\rho h)^{n+1}\\

@@

\\ \\

@@

After the new mass fractions have been computed, the reaction rates are defined as:
\begin{equation}
(\rho\omegadot_k)^{\outp} = \frac{\rho^{\outp} ( X_k^{\outp} - X_k^{\inp})}{\Delta t /2} \enskip,
\end{equation}
and the nuclear energy generation rate is defined as
\begin{equation}
(\rho\Hnuc)^{\outp} = -\sum_k(\rho\omegadot_k)^{\outp} \; q_k \enskip.
\end{equation}
for the case where only strong-mediated reactions are involved.  As
$\Hnuc$ is an output of the reaction network, more general networks
(i.e.\ involving weak interactions) can easily be added.  The enthalpy
update incorporates the external heating, $(\rho\Hext)^{\inp}$, and is
updated by
\begin{equation}
(\rho h)^{\outp} = 
(\rho h)^{\inp} + \frac{\dt}{2} (\rho\Hnuc)^{\outp} + \frac{\dt}{2} (\rho\Hext)^{\inp}\enskip.
\end{equation}

@@

  \item Compute the species edge states, $(\rho X_k)^{(1),\nph,\star}$,
  for the conservative update of $(\rho X_k)$. 
  Here we predict $\rho'$ and $X_k$ to the edges to obtain 
  $\rho^{'(1),\nph,\star}$ and $X_k^{(1),\nph,\star}$ using a second-order
  Taylor expansion in space and time, as described in paper II,
  Appendix A, using $\Vb = \uadvone+w_0^{\nph,\star} \er$.  We use the base
  state density, $\rho_0$, to convert this to an edge state for $\rho$,
\begin{equation}
\rho^{(1),\nph,\star} = 
\rho^{'(1),\nph,\star} + \frac{\rho_0^n + \rho_0^{(2a),\star}}{2} \enskip .
\label{eq:rhoXedgestate}
\end{equation}
  In plane-parallel, we use 4th-order spatial averaging to compute $\rho_0$ on 
  radial faces.  In spherical, we first map $\rho_0$ to Cartesian cell-centers 
  using the mapping in Section \ref{Sec:1D Cell-Centered to Cartesian Cell-Centered}, 
  and then use 2nd-order spatial averaging to compute $\rho_0$ on all faces.

@@

  \item Compute the enthalpy edge state, $(\rho h)^{(1),\nph,\star},$
    for the conservative update of $(\rho h).$  Here we predict $(\rho
    h)'$ to the edges to obtain $(\rho h)^{'(1),\nph,\star}$, 
    using a second-order Taylor expansion in space
    and time, as described in paper II, Appendix A, using $\Vb =
    \uadvone+w_0^{\nph,\star} \er$.  We do not include the reaction
    terms in the enthalpy prediction, since we accounted for them
    already in {\bf React State}.  We use the base state enthalpy,
    $(\rho h)_0$, to convert this to an edge state for $(\rho h)$,

\begin{equation}
(\rho h)^{(1),\nph,\star} = 
(\rho h)^{'(1),\nph,\star} + \frac{(\rho h)_0^n + (\rho h)_0^{(2),\star}}{2}
\enskip .
\end{equation}
  In plane-parallel, we use 4th-order spatial averaging to compute $(\rho h)_0$
  on radial faces.  In spherical, we first map $h_0 \equiv (\rho h)_0/\rho_0$ 
  and $\rho_0$ to Cartesian cell-centers using the mapping in Section 
  \ref{Sec:1D Cell-Centered to Cartesian Cell-Centered}, use 2nd-order spatial 
  averaging to compute $h_0$ and $\rho_0$ on all faces, and then multiply these 
  terms to get $(\rho h)_0$ on all faces.

@@

\begin{equation}
\rho^{(1),\nph} = \rho^{'(1),\nph} + \frac{\rho_0^n + \rho_0^{(2a)}}{2}\enskip .
\end{equation}
  In plane-parallel, we use 4th-order spatial averaging to compute $\rho_0$ on 
  radial faces.  In spherical, we first map $\rho_0$ to Cartesian cell-centers 
  using the mapping in Section \ref{Sec:1D Cell-Centered to Cartesian Cell-Centered}, 
  and then use 2nd-order spatial averaging to compute $\rho_0$ on all faces.

@@

  \item Compute the enthalpy edge state, $(\rho h)^{(1),\nph}$, for
    the conservative update of $(\rho h)$. Here we predict $(\rho h)'$
    to the edges to obtain $(\rho h)^{'(1),\nph}$, 
    using a second-order Taylor expansion in space and
    time, as described in paper II, Appendix A, using $\Vb =
    \uadvtwo+w_0^{\nph} \er$.  Again, we do not include the reaction
    terms in the enthalpy prediction since we accounted for them
    already in {\bf React State}. We use the base state enthalpy,
    $(\rho h)_0$, to convert these to edge states for $(\rho h)$,
\begin{equation}
(\rho h)^{(1),\nph} = 
(\rho h)^{'(1),\nph} + \frac{(\rho h)_0^n + (\rho h)_0^{(2)}}{2} \enskip .
\end{equation}
  In plane-parallel, we use 4th-order spatial averaging to compute $(\rho h)_0$
  on radial faces.  In spherical, we first map $h_0 \equiv (\rho h)_0/\rho_0$ 
  and $\rho_0$ to Cartesian cell-centers using the mapping in Section 
  \ref{Sec:1D Cell-Centered to Cartesian Cell-Centered}, use 2nd-order spatial 
  averaging to compute $h_0$ and $\rho_0$ on all faces, and then multiply these 
  terms to get $(\rho h)_0$ on all faces.

### 2009-01-21 docs/radial_base/radial_base.tex
make the radial base figure work whether we build the main document or build just the radial base section alone
proper reference to spherical averaging
fixed some typos, added more detail about mapping and where it happens in the algorithm

As there is no alignment between the 3-d Cartesian grid and the 1-d
spherical base state grid, we are free to choose their grid spacings
independently.  Experimentation has shown that the best choice results
from picking the 3-d Cartesian grid spacing, $\Delta x$ to be coarser
than the 1-d spherical base state spacing, $\Delta r$.
A typical simulation may use $5 \Delta r = \Delta x$.
We call the procedure to go from 1-d to 3-d {\tt fill\_3d} and the
complementary procedure that takes us from 3-d to 1-d {\tt average}.
\section{\tt fill\_3d}
There are three different {\tt fill\_3d} modes:
\begin{enumerate}
\item The 1D array is bin-centered, and we map to Cartesian cell-centers.
\item The 1D array is edge-centered, and we map to Cartesian cell-centers.
\item The 1D array is edge-centered, and we map to Cartesian edges.  
\end{enumerate}
\subsection{1D Cell-Centered to Cartesian Cell-Centered}
This mapping is done in the subroutine {\tt put\_1d\_array\_on\_cart}.
Figure \ref{fig:mapping} shows the Cartesian grid with the spherical
base state overlayed (for simplicity, the figure is drawn in 2-d).  
We compute the radius of the Cartesian cell center from the center of the star,
\begin{equation}
r = \sqrt{(x_i - x_c)^2 + (y_j - y_c)^2 + (z_k - z_c)^2}
\end{equation}
where $(x_c, y_c, z_c)$ are the coordinates of the center of the star.
We use this radius to find the corresponding radial bin as $n = \mathtt{int}(r
/ \Delta r)$ (here, our convention is to use 0-based indexing for the
base state).  We can then initialize a Cartesian cell quantity $q$ from its
corresponding base state quantity, $q_0$ as $q_{i,j,k} = q_{0,n}$.  In essence,
we are doing a piecewise constant interpolation of the radial array onto the 
Cartesian grid.  If the output is a vector quantity, rather than a scalar, we
multiply the result by the unit vector in the normal direction.
\subsection{1D Edge-Centered to Cartesian Cell-Centered}
This mapping is done in the subroutine {\tt put\_1d\_array\_on\_cart}.
Again, for each Cartesian cell-center, we compute the corresponding radial
bin.  We linearly interpolate from the two nearest radial edge values 
to get the Cartesian cell-centered value.  In essence,
we are doing a piecewise linear interpolation of the radial array onto the 
Cartesian grid.  If the output is a vector quantity, rather than a scalar, we
multiply the result by the unit vector in the normal direction.
\subsection{1D Edge-Centered to Cartesian Edges}
This mapping is done in the subroutine {\tt put\_1d\_array\_on\_cart}.
This case is only used to create {\tt w0mac} from {\tt w0} in the 
subroutine, {\tt put\_w0\_on\_edges}.  There are several variations 
of the interpolation, controlled by {\tt w0mac\_interp\_type}.  We 
currently use the default, {\tt w0mac\_interp\_type} = 1.  We first 
map $w_0$ onto Cartesian cell-centers, and multiply the result by 
the unit vector in the normal direction.  Then, for each Cartesian 
face, we obtain the normal velocity by averaging the corresponding 
velocities from the Cartesian cell-centers adjacent to the face.
\section{\tt average}
For the {\tt average} process, we first create a coarse 1D radial array 
with $\Delta x = \Delta r$.  Then, we compute the radius of the Cartesian cell 
center, and compute the index, $n$, of the corresponding coarse radial 
bin.  $q_{0,n}$ is then the average of all the $q_{i,j,k}$ whose 
Cartesian cell center maps into coarse radial bin $n$.  Then, we 
interpolate the coarse radial array onto the finer radial array.

@@

\begin{figure}[tpb]
\begin{center}
\epsffile{./radial_base/mapping.eps}
\caption{\label{fig:mapping} The Cartesian grid and spherical base
state (shown here in 2-d for simplicity).  Here we represent the
spherical base state as concentric shells (blue lines).  Since the
base state is not aligned with the Cartesian grid we use on which
we discretize the star, we need to map between the two configurations.
The `$+$' symbols represent the Cartesian zone centers.  In our
mapping, the zone marked with the red symbol maps into the gray-shaded
radial bin. }
\label{fig:mapping}
\end{center}
\end{figure}

>>>>

As there is no alignment between the 3-d Cartesian grid and the 1-d
spherical base state grid, we are free to choose their grid spacings
independently.  Experimentation has shown that the best choice results
from picking the 3-d Cartesian grid spacing, $\Delta x$ to be coarser
than the 1-d spherical base state spacing, $\Delta r$.
A typical simulation may use $5 \Delta r = \Delta x$.
We call the procedure to go from 1-d to 3-d {\tt fill\_3d} and the
complementary procedure that takes us from 3-d to 1-d {\tt average}.
\section{\tt fill\_3d}
There are three different {\tt fill\_3d} modes:
\begin{enumerate}
\item The 1D array is bin-centered, and we map to Cartesian cell-centers.
\item The 1D array is edge-centered, and we map to Cartesian cell-centers.
\item The 1D array is edge-centered, and we map to Cartesian edges.  
\end{enumerate}
\subsection{1D Cell-Centered to Cartesian Cell-Centered}
\label{Sec:1D Cell-Centered to Cartesian Cell-Centered}
This mapping is done in the subroutine {\tt put\_1d\_array\_on\_cart}.
Figure \ref{fig:mapping} shows the Cartesian grid with the spherical
base state overlayed (for simplicity, the figure is drawn in 2-d).  
We compute the radius of the Cartesian cell center from the center of the star,
\begin{equation}
r = \sqrt{(x_i - x_c)^2 + (y_j - y_c)^2 + (z_k - z_c)^2}
\end{equation}
where $(x_c, y_c, z_c)$ are the coordinates of the center of the star.
We use this radius to find the corresponding radial bin as $n = \mathtt{int}(r
/ \Delta r)$ (here, our convention is to use 0-based indexing for the
base state).  We can then initialize a Cartesian cell quantity $q$ from its
corresponding base state quantity, $q_0$ as $q_{i,j,k} = q_{0,n}$.  In essence,
we are doing a piecewise constant interpolation of the radial array onto the 
Cartesian grid.  If the output is a vector quantity, rather than a scalar, we
multiply the result by the unit vector in the normal direction.
\subsection{1D Edge-Centered to Cartesian Cell-Centered}
\label{Sec:1D Edge-Centered to Cartesian Cell-Centered}
This mapping is done in the subroutine {\tt put\_1d\_array\_on\_cart}.
Again, for each Cartesian cell-center, we compute the corresponding radial
bin.  We linearly interpolate from the two nearest radial edge values 
to get the Cartesian cell-centered value.  In essence,
we are doing a piecewise linear interpolation of the radial array onto the 
Cartesian grid.  If the output is a vector quantity, rather than a scalar, we
multiply the result by the unit vector in the normal direction.
\subsection{1D Edge-Centered to Cartesian Edges}
\label{Sec:1D Edge-Centered to Cartesian Edges}
This mapping is done in the subroutine {\tt put\_1d\_array\_on\_cart}.
This case is only used to create {\tt w0mac} from {\tt w0} in the 
subroutine, {\tt put\_w0\_on\_edges}.  There are several variations 
of the interpolation, controlled by {\tt w0mac\_interp\_type}.  We 
currently use the default, {\tt w0mac\_interp\_type} = 1.  We first 
map $w_0$ onto Cartesian cell-centers, and multiply the result by 
the unit vector in the normal direction.  Then, for each Cartesian 
face, we obtain the normal velocity by averaging the corresponding 
velocities from the Cartesian cell-centers adjacent to the face.
\section{Avg}\label{Sec:Avg}
For the {\bf Avg} process, we first create a coarse 1D radial array 
with $\Delta x = \Delta r$.  Then, we compute the radius of the Cartesian cell 
center, and compute the index, $n$, of the corresponding coarse radial 
bin.  $q_{0,n}$ is then the average of all the $q_{i,j,k}$ whose 
Cartesian cell center maps into coarse radial bin $n$.  Then, we 
interpolate the coarse radial array onto the finer radial array.

@@

\begin{figure}[tpb]
\begin{center}
\epsffile{\radbasefigpath/mapping.eps}
\caption{\label{fig:mapping} The Cartesian grid and spherical base
state (shown here in 2-d for simplicity).  Here we represent the
spherical base state as concentric shells (blue lines).  Since the
base state is not aligned with the Cartesian grid we use on which
we discretize the star, we need to map between the two configurations.
The `$+$' symbols represent the Cartesian zone centers.  In our
mapping, the zone marked with the red symbol maps into the gray-shaded
radial bin. }
\label{fig:mapping}
\end{center}
\end{figure}
